PWD.KAZ

### 2019-07-16 results/PWD.KAZ.BivData.Rmd
adding updated merged data for the PWD-KAZ analysis

---
title: "PWD.KAZ"
author: "April Peterson"
date: "June 4, 2019"
output: pdf_document
---

@@

setwd("~./MLH1repo/")
#load BivData
load(file="~./MLH1repo/data/MLH1_BivData.RData")

#load MLH1 data
#setwd("C:/Users/April/Desktop/MLH1repo")
load(file="~./MLH1repo/data/MLH1/MLH1_data_setup.RData")

@@

Comparison of PWD and KAZ Biv Data

PWD blah

CZECHII blah

KAZ blah

```{r set.up.data, echo=FALSE}

@@

These are the MLH1 counts

```{r mergeing.meta.data, echo=FALSE}
meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375

#find a way to remove the rows without the right format before setting as as.Date
#trying to remove these the extra rows..uggh
#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]

#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
#i think subseting is messing up the dataframes
meta.data.file99 <- meta.data.file2
meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
#meta.data.file2 <- add_age(meta.data.file2)  #age second
##got it to work ... maybe should save adding it to the merged DF

@@

@@

```{r}
#BivData

@@

@@

```{r}
#distribution of ages

>>>>

---
title: "PWD.KAZ"
author: "April Peterson"
date: "June 4, 2019"
output:
  html_document:
    df_print: paged
---

@@

library(plyr)
library(ggplot2)

setwd("~./MLH1repo/")
#load BivData
#load(file="~./MLH1repo/data/MLH1_BivData.RData")

#load MLH1 data
#setwd("C:/Users/April/Desktop/MLH1repo/") #don't have two
load(file="~./MLH1repo/data/MLH1/MLH1_data_setup_62419.RData")
#MLH1_data_setup_62419.RData
#KAZ_male_data = read.csv("~./data/KAZ_BivData.csv")
meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)

#curBivData_18may19_PWD_m1 = read.csv("~./MLH1repo/results/curate_18may18_PWD_m1_BivData.txt", header = TRUE, sep = '\t')
#curBivData_21may18_KAZ_m2 = read.csv("~./MLH1repo/results/curate_21may18_KAZ_m2_BivData.txt",header = TRUE, sep = '\t')

#PWD set 4 mice, ~100 objs each)
PWD_curate_BivData.full =
  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)
#remove bad segmentation
PWD_curate_BivData <- PWD_curate_BivData.full[PWD_curate_BivData.full$SC.pass == 1,]

#KAZ 3 mice >200 objs each
KAZ_curate_BivData.full =
  read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE)
KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]

#load hand measures, get it to match the auto
PWD_hand_measure_BivData.full =
  read.csv("~./MLH1repo/data/BivData/PWD_male_hand_measures.csv", header = TRUE)

#remove extra empty rows
PWD_hand_measure_BivData.full <- PWD_hand_measure_BivData.full[!(is.na(PWD_hand_measure_BivData.full$fileName) | PWD_hand_measure_BivData.full$fileName == ""), ]

#PWD_hand_measure_BivData.full$Foci1  <- PWD_hand_measure_BivData.full$foci.1.position
#PWD_hand_measure_BivData.full$Foci2 <- PWD_hand_measure_BivData.full$medial.foci.position
#PWD_hand_measure_BivData.full$Foci3 <- PWD_hand_measure_BivData.full$distal.foci.position
#PWD_hand_measure_BivData.full$boxNumber <- PWD_hand_measure_BivData.full$manual.blobject.name

PWD_hand_measure_BivData.full$Obj.ID <- paste(PWD_hand_measure_BivData.full$fileName, PWD_hand_measure_BivData.full$boxNumber, sep = "_")

PWD_hand_measure_BivData.full$centromere_PER_Position <- PWD_hand_measure_BivData.full$centromere_ABS_Position / PWD_hand_measure_BivData.full$chromosomeLength

PWD_hand_measure_BivData.full$SC.pass <- "1"
PWD_hand_measure_BivData.full$foci.pass <- "1"
PWD_hand_measure_BivData.full$curated.1pass.0fail. <- ""
#PWD_hand_measure_BivData.full$curation.notes <- ""
#PWD_hand_measure_BivData.full$hand.foci.count <- ""

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
PWD_hand_measure_BivData.full <- add_mouse(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_category(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_strain(PWD_hand_measure_BivData.full)
PWD_hand_measure_BivData.full <- add_sex(PWD_hand_measure_BivData.full)

#3 columns don't overlap
#avergeRED, numberCrossOvers, 
#manual.blobject.name,

KAZ_hand_measure_BivData.full =
  read.csv("~./MLH1repo/data/BivData/KAZ_male_hand_measures.csv", header = TRUE)

KAZ_hand_measure_BivData.full <- KAZ_hand_measure_BivData.full[!(is.na(KAZ_hand_measure_BivData.full$fileName) | KAZ_hand_measure_BivData.full$fileName == ""), ]

KAZ_hand_measure_BivData.full$boxNumber <- KAZ_hand_measure_BivData.full$manual.blobject.name

KAZ_hand_measure_BivData.full$Obj.ID <- paste(KAZ_hand_measure_BivData.full$fileName, KAZ_hand_measure_BivData.full$boxNumber, sep = "_")

KAZ_hand_measure_BivData.full$SC.pass <- "1"
KAZ_hand_measure_BivData.full$foci.pass <- "1"
KAZ_hand_measure_BivData.full$curated.1pass.0fail. <- ""

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
KAZ_hand_measure_BivData.full <- add_mouse(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_category(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_strain(KAZ_hand_measure_BivData.full)
KAZ_hand_measure_BivData.full <- add_sex(KAZ_hand_measure_BivData.full)

#not sure why I have a distinction of hand_measure and hand_measure.full
#KAZ_hand_measure_BivData <- KAZ_hand_measure_BivData.full[KAZ_hand_measure_BivData.full$SC.pass == 1,]

#manual blobject name -- boxNumber, Obj.ID 
#  20jul15_18may15_PWD_m1_sp1_2.2_blob1
```

@@

```{r, echo=FALSE}
#Curated data
#merge PWD and KAZ BivData

@@

what are the distributions of SC lengths?

```{r cell.level.tables, echo=FALSE}

#make cell level tables
#id the 18-20 biv cells

#estimate which cells can be used to make grouped pools

### SC Length Analysis

Make plots for SC lengths

```{r SC.lengths}
library(ggplot2)
#SC lenghts

#strange number of low chrm lengths
#hist(SC.passed.PWD$chromosomeLength)

SC.PWD <- ggplot(data = PWD_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("PWD SC lengths")
#the SC range is 45 - 150 (for PWD)

SC.KAZ <- ggplot(data = KAZ_curate_BivData, aes(color=fileName,x=mouse, y= chromosomeLength))+geom_jitter()+theme(legend.position="none")+ylim(c(35,150))+ggtitle("KAZ SC lengths")
#can I add the means?

SC.PWD
SC.KAZ
```

```{r cell.tables}
#add total COs
KAZ_BivData_cells <- ddply(KAZ_curate_BivData, c("fileName"), summarise,

                    boxs.IDd = max(boxNumber),
                    nboxes_passed = length(boxNumber),
                    pass_rate = nboxes_passed/boxs.IDd,

                    totalCO = sum(numberCrossOvers),     
                    avCO.per.CHRM = mean(numberCrossOvers),
                    varCO.per.CHRM = var(numberCrossOvers),

                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
                    varSC.intensity = var(avergeRED, na.rm = TRUE),

                    avSC.length = mean(chromosomeLength),
                    varSC.length = var(chromosomeLength),

                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
                    varCOdensity = var(chromosomeLength /numberCrossOvers)
                            #av CO density
                    #change the mean list for column numbers

)

cells2curate <- KAZ_BivData_cells[KAZ_BivData_cells$pass_rate > .7,]

write.table(cells2curate, "~./MLH1repo/results/KAZ.cell.curate.csv", sep=",", row.names = FALSE)

#PWD version
#add total COs
PWD_BivData_cells <- ddply(PWD_curate_BivData, c("fileName"), summarise,

                    boxs.IDd = max(boxNumber),
                    nboxes_passed = length(boxNumber),
                    pass_rate = nboxes_passed/boxs.IDd,

                    totalCO = sum(numberCrossOvers),     
                    avCO.per.CHRM = mean(numberCrossOvers),
                    varCO.per.CHRM = var(numberCrossOvers),

                    avSC.intensity = mean(avergeRED, na.rm = TRUE),
                    varSC.intensity = var(avergeRED, na.rm = TRUE),

                    avSC.length = mean(chromosomeLength),
                    varSC.length = var(chromosomeLength),

                    av.COdensity = mean(chromosomeLength /numberCrossOvers),
                    varCOdensity = var(chromosomeLength /numberCrossOvers)
                            #av CO density
                    #change the mean list for column numbers

)

PWDcells2curate <- PWD_BivData_cells[PWD_BivData_cells$pass_rate > .65,]

write.table(PWDcells2curate, "~./MLH1repo/results/PWD.cell.curate.csv", sep=",", row.names = FALSE)

From the cell level tables -- make lists of cells with ~75% of bivalents quantified. These will be manually verified and the missing chrms will be measured and supplemented.

```{r PWD.merge.manual.auto, echo=FALSE}
#merging manual measures and auto measures
#KAZ_hand_measure_BivData
#PWD_hand_measure_BivData

#.1 match the colnames,  #merge, partial match
#colnames(PWD_curate_BivData)
#colnames(PWD_hand_measure_BivData.full)

#merg4.PWD_curate_BivData <-    #remove "avergeRED", numberCrossOvers
#merg4.PWD_hand_measure_BivData.full <- #manual.blobject.name

#auto.curate
auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
new.DF <- PWD_curate_BivData[ , auto.c_names, drop = TRUE]

#manual
manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position","SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
new.DF2 <- PWD_hand_measure_BivData.full[ , manual_names, drop = TRUE]

#2. merge the 2 DFs
PWD.auto.man_BivData <- rbind(new.DF, new.DF2)

#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
#make better variable name
PWD.auto.man_BivData_cells <- ddply(PWD.auto.man_BivData, c("fileName"), summarise,
                    number = length(Obj.ID),
                    #make the rank or
                     maxChrm_length = max(chromosomeLength),
                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]

)

#Isolate cells with all autosomes measured
PWDfull.cell.list <- PWD.auto.man_BivData_cells[PWD.auto.man_BivData_cells$number > 17,]
PWDfull.cell.list <- PWDfull.cell.list[PWDfull.cell.list$number < 21,]#remove the too many cell

#this gives the rank to each biv within each cell  
PWD.Bivs.w.rank <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
#there are instances of ties,

#this does something weird
#PWD.auto.man_BivData$rank1 <- ddply(PWD.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)

PWD.auto.man_BivData <- PWD.auto.man_BivData[ -c(19:25) ]

#pull all biv data for the full cell data
PWD_whole.cell.BivData <- PWD.Bivs.w.rank[PWD.Bivs.w.rank$fileName %in% PWDfull.cell.list$fileName,]
#ok these should be full cells
hist(hh2$rank, breaks =21)
#there are 2x as many 0's bc 1 and 0 are binned together
#update all hand.foci.counts
#y <- droplevels(y)

length.by.class <- ggplot(data=hh2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=hand.foci.count, shape=mouse))+ggtitle("PWD full cell SC lengths")
#most cells are from 1 mouse

DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName %in% full.cell.list$fileName,]
DF.w.rank2 <- DF.w.rank[DF.w.rank$fileName != "20jul15_18may15_PWD_m1_sp1_17_rev",]

qq <- ggplot(data=DF.w.rank2, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=rank, shape=mouse))+ggtitle("PWD full cell SC lengths")
#most cells are from 1 mouse

#take the shortest 5? (lower 25%)
#take the shortest 3? (seems to be natural cutoff, lowest 3 are similar)

#bin.long.PWD <- DF.w.rank[DF.w.rank$rank]
#can't do long yet, bc the coding for XY is inconsistant!

bin.short.PWD <- DF.w.rank2[DF.w.rank2$rank < 5,]
bin.short.PWD <- bin.short.PWD[bin.short.PWD$fileName %in% full.cell.list$fileName,]

kk <- ggplot(data=bin.short.PWD, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color =Foci1, shape=mouse))+ggtitle("short bins")

#short bin patterns vs long bins / total?
#eventually I want to compare NORM FOCI POSITIONS patterns of these bins across PWD and KAZ
#merging is a little too confusing... 
#just parse down and do rbind?
m1 <- merge(PWD_curate_BivData, PWD_hand_measure_BivData.full, by.x = Obj.ID)

#KAZ_curate_BivData
#KAZ_hand_measure_BivData.full

hist(hh2$rank, breaks =21)
```

Asigned rank to bivalent measures from the whole.cell dataset. I should try to use these measures to see if I can guess the ranks of 'loose' bivalent measures.
I also noted that the shortest 3 seem to cluster in PWD. Maybe I should try doing cluster analysis on these. 
I should also add rbar values to these measures. (I might have to reformat the function)

```{r KAZ.manual.merge}
#KAZ_hand_measure_BivData
#PWD_hand_measure_BivData
#KAZ_curate_BivData <- KAZ_curate_BivData.full[KAZ_curate_BivData.full$SC.pass == 1,]

#1 match the colnames,  #merge, partial match
colnames(KAZ_curate_BivData)
colnames(KAZ_hand_measure_BivData.full)

KAZ_hand_measure_BivData.full$centromere_PER_Position <- KAZ_hand_measure_BivData.full$centromere_ABS_Position / KAZ_hand_measure_BivData.full$chromosomeLength

KAZ_hand_measure_BivData.full$curation.notes <- KAZ_hand_measure_BivData.full$notes

#auto; avergeRED, numberCrossOvers, 20
#manual; 15

#auto.curate
KAZ.auto.c_names <- c("fileName","centromere_ABS_Position","centromere_PER_Position",  "chromosomeLength","boxNumber","Obj.ID","SC.pass","foci.pass","curated.1pass.0fail.","curation.notes","hand.foci.count","Foci1","Foci2","Foci3" ,"mouse", "strain","sex","category")
new.DF <- KAZ_curate_BivData[ , KAZ.auto.c_names, drop = TRUE]

#manual
KAZ.manual_names <- c("fileName","chromosomeLength","centromere_ABS_Position","boxNumber","Obj.ID","Foci1","Foci2","Foci3","centromere_PER_Position", "SC.pass","foci.pass","curated.1pass.0fail.", "mouse", "category","strain","sex","curation.notes", "hand.foci.count")
new.DF2 <- KAZ_hand_measure_BivData.full[ , KAZ.manual_names, drop = TRUE]

#2. merge the 2 DFs
KAZ.auto.man_BivData <- rbind(new.DF, new.DF2)

#some cleanup may be required for the manual data; remove all XY, mark chrm classes for all 
#make better variable name
KAZ.auto.man_BivData_cells <- ddply(KAZ.auto.man_BivData, c("fileName"), summarise,
                    number = length(Obj.ID),
                    #make the rank or
                     maxChrm_length = max(chromosomeLength),
                    maxChrm_boxNumber = boxNumber[which.max(chromosomeLength)]

)

#Isolate cells with all autosomes measured
KAZfull.cell.list <- KAZ.auto.man_BivData_cells[KAZ.auto.man_BivData_cells$number > 17,]
KAZfull.cell.list <- KAZfull.cell.list[KAZfull.cell.list$number < 21,]#remove the too many cell

#this gives the rank to each biv within each cell  
KAZ.Bivs.w.rank <- ddply(KAZ.auto.man_BivData, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
#there are instances of ties,

KAZ_whole.cell.BivData <- KAZ.Bivs.w.rank[KAZ.Bivs.w.rank$fileName %in% KAZfull.cell.list$fileName,]
#ok these should be full cells
hist(KAZ_whole.cell.BivData$rank, breaks =25)
KAZ_whole.cell.BivData <- droplevels(KAZ_whole.cell.BivData)
table(KAZ_whole.cell.BivData$fileName, KAZ_whole.cell.BivData$rank)

KAZ.length.by.class <- ggplot(data=KAZ_whole.cell.BivData, aes(x=fileName,y=chromosomeLength))+geom_jitter(width = 0.25, aes(color=as.factor(hand.foci.count), shape=mouse))+ggtitle("PWD full cell SC lengths")

KAz have more 0CO. Most of the NA's for CO count are from auto- where I was whisy-washy during curation.

```{r stat.test.SC.length, echo=FALSE}
#can't cleanly merge the curated data since the rows and headers are different
merged.cur <- rbind(curBivData_PWD_m1, curBivData_KAZ_m2)

SCs <- ggplot(data = , aes(y, x=mice))+geom_point

SC.test <- t.test()

```{r IDF, echo=FALSE}

#isolate the 2COs
#total Bivdata
KAZ.2COs <- KAZ.auto.man_BivData[KAZ.auto.man_BivData$hand.foci.count == 2,]
PWD.2COs <- PWD.auto.man_BivData[PWD.auto.man_BivData$hand.foci.count == 2,]

all.2CO <- rbind(KAZ.2COs,PWD.2COs)
#all.2CO <- ddply(all.2CO, .(fileName), transform, rank = rank(chromosomeLength,ties.method ="min")-1)
#all.2CO <- all.2CO[ -c(24) ]

all.2CO$Foci1.PER <- all.2CO$Foci1 / all.2CO$chromosomeLength
all.2CO$Foci2.PER <- all.2CO$Foci2 / all.2CO$chromosomeLength
all.2CO$Foci3.PER <- all.2CO$Foci3 / all.2CO$chromosomeLength

#call IFD function on this DF
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
all.2CO <- add_IFD(all.2CO)
all.2CO <- droplevels(all.2CO)
all.2CO <- all.2CO[!(is.na(all.2CO$fileName) | all.2CO$fileName == ""), ]

all.2CO$mouse <- droplevels(all.2CO$mouse)

IFDs.ABS <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("raw IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 90, hjust = 1) )

IFDs.PER <- ggplot(data=all.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(fileName) ))+ggtitle("norm IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+theme(legend.position="none")

#whole cell comparison (use rank)
#KAZ_whole.cell.BivData, #PWD_whole.cell.BivData

#limit to mice with most observations

whole.cell.2CO <- rbind(KAZ_whole.cell.BivData,PWD_whole.cell.BivData)

whole.cell.2CO$Foci1.PER <- whole.cell.2CO$Foci1 / whole.cell.2CO$chromosomeLength
whole.cell.2CO$Foci2.PER <- whole.cell.2CO$Foci2 / whole.cell.2CO$chromosomeLength
whole.cell.2CO$Foci3.PER <- whole.cell.2CO$Foci3 / whole.cell.2CO$chromosomeLength

#call IFD function on this DF
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
whole.cell.2CO <- add_IFD(whole.cell.2CO)
whole.cell.2CO <- droplevels(whole.cell.2CO)
whole.cell.2CO <- whole.cell.2CO[!(is.na(whole.cell.2CO$fileName) | whole.cell.2CO$fileName == ""), ]

IFDs.ABS.whole.cell <- ggplot(data=whole.cell.2CO, aes(x=mouse,y=IFD1_ABS))+geom_jitter(width = 0.25, aes(color=as.factor(rank) ))+ggtitle("Whole cell IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )

IFDs.PER.whole.cell <- ggplot(data=whole.cell.2CO, aes(x=mouse,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(rank) ))+ggtitle("Whole Cell Norm IFD Distributions")+geom_boxplot(alpha = 0.6)+facet_wrap(~strain)+theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )

```{r IFD.SC, echo=FALSE}
#compare the realtionship of SC length and

values.ABS.all <- all.2CO %>% 
    group_by(strain) %>% 
    do({
      mod = lm(IFD1_ABS ~ chromosomeLength, data = .)
      data.frame(Intercept = coef(mod)[1],
                 Slope = coef(mod)[2])
    })

SC.IFD <- ggplot(data=all.2CO, aes(x=chromosomeLength,y=IFD1_ABS))+geom_point( aes(color=as.factor(mouse) ))+
  ggtitle("SC by IFD length")+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )+ 
  geom_smooth(method="lm")#+
  #geom_text(values.ABS.all, aes(x = 75, y = 100, label = n), parse = TRUE, inherit.aes=FALSE)

SC.IFD_PER <- ggplot(data=all.2CO, aes(x=chromosomeLength,y=IFD1_PER))+geom_jitter(width = 0.25, aes(color=as.factor(mouse) ))+
  geom_abline(intercept = 0.25)+
    geom_smooth(method="lm")+
  ggtitle("SC by nrm IFD length")+facet_wrap(~strain)+  theme(legend.position="none", axis.text.x = element_text(angle = 45, hjust = 1) )

```{r mergeing.meta.data, echo=FALSE}
#meta.data.file2 = read.csv("~./MLH1repo/data/MouseMetaData61119.csv", header = TRUE)
#meta.data.file2 = read.csv("~./MLH1repo/data/Mouse_MetaData_6.12.19.csv", header = TRUE)

meta.data.file2 <- meta.data.file2[!(is.na(meta.data.file2$mouse)|meta.data.file2$mouse==""),] #375

#find a way to remove the rows without the right format before setting as as.Date
#trying to remove these the extra rows..uggh
#yy <- meta.data.file2[subset(meta.data.file2$DOB, !( (meta.data.file2$DOB == "emb") | (meta.data.file2$DOB == "arr") | (meta.data.file2$DOB == "e16.5")|(meta.data.file2$DOB == "march") ) ),]
#yy <- yy[subset(yy$DOB, !(yy$DOB == "arr")),]
#yy <- yy[subset(yy$DOB, !(yy$DOB == "e1")),]
#i think subseting is messing up the dataframes

meta.data.file2$DOB <- as.Date(meta.data.file2$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,

#source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
#meta.data.file2 <- add_euth_date(meta.data.file2) #euthdate 1st
#meta.data.file2 <- add_age(meta.data.file2)  #age second
##got it to work ... maybe should save adding it to the merged DF
meta.data.file2 <- add_category(meta.data.file2)
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")
meta.data.file2 <- add_euth_date(meta.data.file2)
meta.data.file2 <- add_age(meta.data.file2)

pwd.meta <- meta.data.file2[meta.data.file2$category == "PWD male",]

#pwd.meta <- add_euth_date(pwd.meta)
#pwd.meta <- add_age(pwd.meta)

kaz.meta <- meta.data.file2[meta.data.file2$category == "KAZ male",]

#kaz.meta <- add_mouse(kaz.meta)
#kaz.meta <- add_euth_date(kaz.meta)
county =1
for(i in kaz.meta$mouse){
    templist= strsplit(i, split="_")[[1]]
    euth <- templist[1]
    #figure out date conversion
    kaz.meta$euth.date2[county] <- as.character(euth)

    #kaz.meta$euth.date[county] <- as.Date(kaz.meta$euth.date[county], format='%d%b%y')
    county= county +1
}

kaz.meta$euth.date <- as.Date(kaz.meta$euth.date2, format='%d%b%y')

kaz.meta <- add_age(kaz.meta)

@@

```
##########################
# MELTED CHRM1 POSITIONS #
##########################

```{r PWD.melt.foci.data, echo=FALSE}
library(reshape2)

newPWD_curate_BivData =
  read.csv("~./MLH1repo/data/BivData/Merged_Curate_PWD_BivData.csv", header = TRUE)

#update these dataframes!
#isolate the passing foci 
curBivData_PWD.good.foci  <- newPWD_curate_BivData[newPWD_curate_BivData$foci.pass == 1,]

#create normalized foci values!
curBivData_PWD.good.foci$nrm.F1 <- curBivData_PWD.good.foci$Foci1 / curBivData_PWD.good.foci$chromosomeLength
curBivData_PWD.good.foci$nrm.F2 <- curBivData_PWD.good.foci$Foci2 / curBivData_PWD.good.foci$chromosomeLength
curBivData_PWD.good.foci$nrm.F3 <- curBivData_PWD.good.foci$Foci3 / curBivData_PWD.good.foci$chromosomeLength

#melting the DataFrame is needed to ploting the normalized foci positions

#do not include cols for IFD, or ABS.foci
curBivData_PWD.melt <- melt( curBivData_PWD.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","nrm.F1","nrm.F2","nrm.F3",  "mouse", "category",   "strain" , "sex")] ,

                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","mouse", "category",   "strain" , "sex"))

colnames(curBivData_PWD.melt) <- c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count", "mouse", "category",   "strain" , "sex",
                                       #name for the 2 melted cols
                                      "Foci.Type.nrm", "Foci.nrm.pos")

#remove NAs
curBivData_PWD.melt <- curBivData_PWD.melt[!(is.na(curBivData_PWD.melt$Foci.nrm.pos) | curBivData_PWD.melt$Foci.nrm.pos == ""), ]

table(newPWD_curate_BivData$fileName)
table(newPWD_curate_BivData$mouse) #4 mice, ~100 each mouse

@@

```{r curated, echo=FALSE}
curBivData_KAZ.good.foci <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
#curated objects
KAZ_m2_curate_bivData =
  read.csv("~./MLH1repo/data/BivData/curated_21may18_KAZ_m2_BivData.csv", header = TRUE)

KAZ_m2_curate_bivData$nrm.F1.pos <- KAZ_m2_curate_bivData$Foci1.pos / KAZ_m2_curate_bivData$chromosomeLength
KAZ_m2_curate_bivData$nrm.F2.pos <- KAZ_m2_curate_bivData$Foci2.pos / KAZ_m2_curate_bivData$chromosomeLength
KAZ_m2_curate_bivData$nrm.F3.pos <-   KAZ_m2_curate_bivData$Foci3.pos / KAZ_m2_curate_bivData$chromosomeLength

#remove those that don't pass
KAZ_m2_curate_bivData_SC_pass <- KAZ_m2_curate_bivData[KAZ_m2_curate_bivData$SC.pass == 1,]
#what else...

table(KAZ_m2_curate_bivData_SC_pass$fileName)

#melt the foci positions, best thing to compare for cells that don't have all the bivalents
#ggplots

show.plot <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=chromosomeLength, x=fileName))+geom_point()

show.plot2 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=Foci1.pos, x=hand.foci.count))+geom_jitter()

show.plot99 <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(y=nrm.F1.pos, x=hand.foci.count))+geom_jitter()+
  ggtitle("Normalized foci positions")

count.hist <- ggplot(data =KAZ_m2_curate_bivData_SC_pass, aes(hand.foci.count))+geom_histogram()

```{r BivData.set, echo=FALSE}
#goal load and set up of the BivData Dataframe

#BivData
# load more bivData
KAZ_bivData = read.csv("~./MLH1repo/data/KAZ_BivData.csv", header = TRUE)

P_fileNames <- KAZ_bivData$fileName[(grep('p_rev', KAZ_bivData$fileName))]
length(P_fileNames)#187

KAZ_bivData <- KAZ_bivData[!(KAZ_bivData$fileName %in% P_fileNames), ]#3642

source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

KAZ_bivData <- add_mouse(KAZ_bivData)
KAZ_bivData <- add_strain(KAZ_bivData)
KAZ_bivData <- add_sex(KAZ_bivData)
KAZ_bivData <- add_category(KAZ_bivData)

#dup_list= c()
#for(i in KAZ_bivData$fileName){  #change the file name header,  
#  # print(i)
#  ifelse(grepl("\\.[0-9]_rev", i),
#         (dup_list = c(dup_list, i)), NA)
#}
#MultiCells = c()
#for(j in dup_list){
#  stringLIST = strsplit(j, split="\\.")[[1]]
#  new_string = paste(stringLIST[[1]][1], "rev", sep="_")
#  MultiCells = c(MultiCells, new_string)
#}
#MultiCells = unique(MultiCells)
#multi_cell_num = length(MultiCells)

#test this part of the function
#KAZ_bivData <- KAZ_bivData[!(KAZ_bivData$fileName %in% MultiCells), ]#35 removed from KAZ data

@@

#merge PWD and KAZ (WHAT ABOUT THE foci?)
#melt the BivData?
```

```{r curatedBivData}

curBivData_KAZ = read.csv("~./MLH1repo/data/BivData/curate_KAZ_BivData.csv", header = TRUE, sep = ',')
#curBivData_PWD = read.csv("~./MLH1repo/data/BivData/curate_PWD_BivData.csv", header = TRUE, sep = ',')

#make normalized

curBivData_KAZ$nrm.F1.pos <- curBivData_KAZ$Foci1 / curBivData_KAZ$chromosomeLength
curBivData_KAZ$nrm.F2.pos <- curBivData_KAZ$Foci2 / curBivData_KAZ$chromosomeLength
curBivData_KAZ$nrm.F3.pos <-   curBivData_KAZ$Foci3 / curBivData_KAZ$chromosomeLength

curBivData_PWD$nrm.F1.pos <- curBivData_PWD$Foci1 / curBivData_PWD$chromosomeLength
curBivData_PWD$nrm.F2.pos <- curBivData_PWD$Foci2 / curBivData_PWD$chromosomeLength
curBivData_PWD$nrm.F3.pos <-   curBivData_PWD$Foci3 / curBivData_PWD$chromosomeLength

#table / what are the numbers of observations
table(curBivData_PWD$fileName)
#biv measures per cell

SC.passed.PWD <- curBivData_PWD[curBivData_PWD$SC.pass == 1,]#299 -> 208
table(SC.passed.PWD$fileName)

SC.passed.KAZ <- curBivData_KAZ[curBivData_KAZ$SC.pass == 1,]#299 -> 208

@@

```{r meltKAZ, echo=FALSE}

#melt KAZ
#isolate the passing foci 
curBivData_KAZ.good.foci  <- curBivData_KAZ[curBivData_KAZ$foci.pass == 1,]
#colnames(curBivData_KAZ)

#melting the DataFrame is needed to ploting the normalized foci positions

#do not include cols for IFD, or ABS.foci
curBivData_KAZ.melt <- melt( curBivData_KAZ.good.foci[,c("fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","nrm.F1.pos","nrm.F2.pos","nrm.F3.pos",  "mouse", "category",   "strain" , "sex")] ,

                                #nrm.F1, nrm.F2, nrm.F3 left out to be melted
              id=c( "fileName", "chromosomeLength", "centromere_ABS_Position", "centromere_PER_Position", "boxNumber","Obj.ID",  "SC.pass","foci.pass","curated.1pass.0fail.", "curation.notes",
    "hand.foci.count","mouse", "category",   "strain" , "sex"))
